---
title: "R/Bioconductor-powered Team Data Science"
author: "Leonardo Collado-Torres"
site: bookdown::bookdown_site
apple-touch-sin: "icon_192.png"
apple-touch-icon-size: 192
favicon: "icon_32.png"
github-repo: "lcolladotor/bioc_team_ds"
documentclass: book
output:
  bookdown::gitbook:
    includes:
      after_body: map_and_ga.html
    css: "style.css"
    
---

# Overview {-}

**R/Bioconductor-powered Team Data Science** by [Leonardo Collado Torres](http://lcolladotor.github.io) and team members.

With lessons learned from former mentors [Jeffrey T. Leek](https://github.com/jtleek) and [Andrew E. Jaffe](https://github.com/andrewejaffe).

## Team members {-}

* [Abby Spangler](https://github.com/abspangler13)
* [Arta Seyedian](https://github.com/aseyedia)
* [Brenda Pardo](https://github.com/bpardo99)
* [Joshua M. Stolz](https://github.com/joshstolz)
* [Louise Huuki](https://github.com/lahuuki)
* [Nicholas J. Eagles](https://github.com/Nick-Eagles)

### Alumni {-}

* [Amy Peterson](https://github.com/Ashkaunr)
* [Ashkaun Razmara](https://github.com/amy-peterson)

## DSgs-guide team members {-}

* [Arta Seyedian](https://github.com/aseyedia)
* [Joshua M. Stolz](https://github.com/joshstolz)
* [Louise Huuki](https://github.com/lahuuki)
* [Nicholas J. Eagles](https://github.com/Nick-Eagles)
* [Madhavi Tippani](https://github.com/madhavitippani)

_Check [Become a DSgs guide](#become-a-dsgs-guide) to join the DSgs-guide team_.

## Acknowledgements {-}

The contents of this book have been refined thanks to feedback from:

* [Keri Martinowich](https://github.com/kmartinow)
* [Kristen R. Maynard](https://github.com/kmaynard12)

This book was last updated on `r Sys.time()`.

## R session information {-}

Details on the R version used for making this book. The source code is available at [`lcolladotor/bioc_team_ds`](https://github.com/lcolladotor/bioc_team_ds).

```{r session_info_chunk, eval = FALSE}
## Load the package at the top of your script
library("sessioninfo")

## Reproducibility information
print('Reproducibility information:')
Sys.time()
proc.time()
options(width = 120)
session_info()
```

```{r session_info, echo = FALSE}
<<session_info_chunk>>
```

